---
title: "Gf_restingState_lasso"
output: html_document
date: '2022-08-17'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(dplyr)
library(magrittr)
library(ggplot2)
library(ggthemes)
library(ppcor)
library(reshape2)
#library(gglasso)
library(glmnet)
library(ggsci)
library(viridis)
library(ggExtra)
#library(kableExtra)
library(xtable)
#library(data.table)
library(sjPlot)
```

## Prediction of Individual Fluid Cognition Scores from Resting-State Neuroimaging Data

This is an analysis pipeline for the prediction of individual measures of a composite fluid cognition score from resting-state functional connectivity data, using the Human Connectome Project dataset. Lasso regression is used to identify the FC features that best predict an individual's composite fluid cognition score.

# Load the fluid cognition scores

```{r}

HCPbehDF = read.csv('/home/ausmanpa/gp/BCB-Modeling/Gf_FC_reg/data/HCP_beh.csv',header=TRUE)

# currently only have data for a subset of HCP participants; would like to expand to full cohort
current_participants = dir('/home/ausmanpa/gp/ACTR-WM/rsfMRI/data')

HCP_Gf = HCPbehDF %>% dplyr::select(Subject,CogFluidComp_Unadj) %>% mutate(Subject=paste0('sub-',Subject)) %>% filter(Subject %in% current_participants)


```


# Plot the histogram and Q-Q plot of fluid cognition scores, test for normality

```{r}
ggplot(HCP_Gf, aes(x=CogFluidComp_Unadj)) +
  geom_histogram(bins=30, col="white", alpha=0.5) +
  geom_vline(xintercept = mean(HCP_Gf$CogFluidComp_Unadj),
             linetype = "dashed") +
  xlab(expression("Fluid Cognition Composite score"))+
  ylab("Number of Participants") +
  ggtitle("Distribution of Fluid Cognition Composite scores") +
  theme_pander() +
  theme(legend.position = "none")


```


```{r}

# http://www.sthda.com/english/wiki/normality-test-in-r

normTest = shapiro.test(HCP_Gf$CogFluidComp_Unadj)

ggplot(HCP_Gf, aes(sample=CogFluidComp_Unadj)) +
  stat_qq() +
  stat_qq_line() +
  annotate('text',x=2, y=90, label=paste0(normTest$method,'\nW = ',round(normTest$statistic,2),'\np = ',round(normTest$p.value,2)))

```

## Load and transform the resting-state data for each participant

First, let's load the Power 2011 region database. This will be used as an "atlas" throughout, to guide the development of the regions.

```{r}
power2011 <- read_csv('/home/ausmanpa/gp/BCB-Modeling/Gf_FC_reg/data/power_2011.csv', 
                      col_types = cols(ROI=col_double(),
                                       X = col_double(),
                                       Y = col_double(),
                                       Z = col_double(),
                                       Network = col_double(),
                                       Color = col_character(),
                                       NetworkName = col_character())) %>%
  dplyr::select(ROI, X, Y, Z, Network, Color, NetworkName)
```

Now create the functional connectivity matrix for every single subject in the dataset. This step can be skipped by setting the `CREATE_FC` variable to `FALSE`.

```{r}
CREATE_FC = FALSE

if (CREATE_FC) {
  for (sub in dir('/home/ausmanpa/gp/ACTR-WM/rsfMRI/data')) {
    roidata = NULL
    for (roi in power2011$ROI) {
      network = power2011 %>%
        filter(ROI == roi) %>%
        dplyr::select(Network) %>%
        as.numeric()
      
      network_name = power2011 %>%
        filter(ROI == roi) %>%
        dplyr::select(NetworkName) %>%
        as.character()
      
      file_name <- paste("region_", 
                         sprintf("%03d", roi),  
                         "_network_", 
                         sprintf("%02d", max(0, network)), 
                         ".txt",
                         sep="")
      
      mat <- colMeans(read.table(paste(sub, 
                                       "ses-01/func",
                                       "swadrfMRI_timeseries_results",
                                       file_name, sep="/")))
      pc1 <- mat
      table <- tibble(subject = sub,
                      scan = 1:1200,
                      timeseries = pc1,
                      roi = roi,
                      network = network,
                      network_name = network_name)
      if (is.null(roidata)) {
        roidata <- table
      } else {
        roidata %<>% bind_rows(table)
      }
      print(paste("sub", sub, "roi", roi))

    }
    # Pivot long data format into wide data 
    wroidata <- roidata %>% pivot_wider(id_cols = scan, 
                                        names_from = roi, 
                                        values_from = timeseries)
    X  <- as.matrix(wroidata[,2:265])
    PR <- pcor(X)$estimate
    R  <- cor(X)
    
    # Generate matrices:
    
    # The partial correlation matrix
    long_pr <- melt(PR)
    #pdf(paste(sub, "fc_pcorr.pdf", sep="/"))
    # ggplot(long_pr, aes(x=Var1, y=Var2)) +
    #   geom_raster(aes(fill=value)) +
    #   scale_fill_gradient2(limits=c(-1,1), 
    #                        low = "blue", 
    #                        high = "red", 
    #                        mid = "white") +
    #   theme_pander() +
    #   ggtitle(paste(sub, ": Functional Connectivity (Partial Correlations)", sep="")) +
    #   xlab("ROIs") +
    #   ylab("ROIs") 
    # #dev.off()
    # ggsave(paste(sub, "fc_pcorr.pdf", sep="/"))
    write.table(PR, col.names = T, 
                row.names = T, 
                file = paste(sub, "PR.txt", sep="/"))
    
    # The standard correlation matrix
    # long_r <- melt(R)
    # #pdf(paste(sub, "fc_corr.pdf", sep="/"))
    # ggplot(long_r, aes(x=Var1, y=Var2)) +
    #   geom_raster(aes(fill=value)) +
    #   scale_fill_gradient2(limits=c(-1,1), low = "blue", high = "red", mid = "white") +
    #   theme_pander() +
    #   ggtitle(paste(sub, ": Functional Connectivity, Standard Correlations)", sep="")) +
    #   xlab("ROIs") +
    #   ylab("ROIs") 
    # #dev.off()
    # ggsave(paste(sub, "fc_corr.pdf", sep="/"))
    write.table(R, col.names = T, 
                row.names = T, 
                file = paste(sub, "R.txt", sep="/"))
  }
}
```

# Perform Lasso regression of individual composite fluid cognition scores on individual resting-state functional connectivity measures

```{r}
fit <- glmnet(y = HCP_Gf$CogFluidComp_Unadj,
              x = X,
              alpha = 1,
              standardize = T
)
```















